\vspace{-15mm}\fontsize{24pt}{10pt}\selectfont\textbf{A Radial Time Projection Chamber for $\alpha$ detection in CLAS at JLab}

### 2017-04-12 eg6-letters/NIM/EG6_RTPC_NIM.tex
NIM paper corrections

% the following line is for submission, including submission to the arXiv!!
%\hspace{5.2in} \mbox{Fermilab-Pub-04/xxx-E}

@@

%\tableofcontents

@@

Measurement of the Deeply Virtual Compton Scattering (DVCS) process
($eH \rightarrow e' H' \gamma$, where $H$ is a hadron) necessitates an upgrade
of this setup.  With a 6 GeV electron beam, the majority of DVCS photons are produced
at very forward angles, where the acceptance of the EC is poor. To extend the detection range,
an inner calorimeter (IC) was built for the E01-113 experiment in 2005.
The IC is constructed from 424 lead-tungstate (PbWO$_{4}$) crystals, covering polar 
angles between 5$^{\circ}$ and 15$^{\circ}$. Each crystal is 16 cm long, corresponding
to 17 radiation lengths. The achieved energy resolution is around 3 to 4$\%$ for photon
energies between 2 GeV and 5 GeV, with an angular resolution between 3 and 5 mrad 
\cite{Hyon-suk}. To protect this detector from the large flux of low energy M{\o}ller 
electrons, a 5~T solenoid was placed around the target to shield the IC. 
In addition, the recoil nuclei of the coherent DVCS on helium was detected 
using a RTPC developed to track low energy nuclear fragments. The solenoid field
serving as the analyzing magnet for the particle tracking in the RTPC. This 
setup was used during a three months experimental run
in 2009 with a longitudinally polarized electron beam of 6.064 GeV incident 
on a gaseous $^{4}$He target.

@@

@@

We determine the drift speed using tracks reconstructed in the RTPC. In figure 
\ref{fig:RTPC_signals}, a typical $^{4}$He track is represented (in green). After 
it causes ionization in the drift region, the released electrons (in black) 
drift to the cylindrical detection plane under the effect of the electric field. The 
electrons released close to the cathode take the most time to reach the readout 
pads, but geometrical symmetry insures they always travel the same 
distance. By identifying the maximum TDC measured, we can infer the drift 
speed of the electrons in the RTPC.\\

@@

We first measure the drift speed along the 200 mm RTPC's length to take into 
account variations in the electric and magnetic field, some of which are known 
(see figure \ref{fig:B_MAP}) some might be due to slight shifts in the 
geometry. We verified this by looking at the TDC profile generated in the 
chamber by all our tracks (see figure \ref{fig:TDC_profile}). The time profile 
of registered hits clearly show the dropping edge expected from the geometrical 
considerations previously mentioned. We define a value named $TDC_{Max/2}$ at 
which the dropping edge passes half the maximum number of hits in the 
histogram, and this value is inversely proportional to the drift speed. The
result in figure \ref{fig:RunNumber_61551_TDCmax_Zslice} shows the dependence of
the drift speed on position along the beam line axis in the chamber.

\begin{figure}[tb]
   \centering
   \includegraphics[scale=0.35]{fig/B_MAP.png}
   \caption[]{Schematic drawing of the solenoid magnetic field vectors inside the 
   RTPC. The $y$-axis is the radial distance from the center of the target, and the
   $x$-axis is the longitudinal position along the RTPC (with its center at $z=0$).} 
\label{fig:B_MAP}
\end{figure}

@@

Due to the non perfect experimental conditions including variations in the 
gas mixture, the drift speed might also change during the month long experimental run.  
Figure \ref{fig:Drift_run_number_1} shows the $TDC_{Max/2}$ values for 
individual runs (approximately 2 hours long). We observe a significant 
variation over time and accounted for these changes in the drift speed in the
track reconstruction. Dealing with such situations is 
possible by adapting the drift speed as a function of 
geometry and time (we did not see any correlation between the effects). These 
functions were extracted for our entire data set and implemented in the 
reconstructions code.

@@

\subsection{Drift Paths Parametrization}

The drift paths are the trajectories the electrons follows in the gas. The 
standard software to calculate drift paths is the MAGBOLTZ program 
~\cite{MAGBOLTZ}, which requires knowledge of the detector's 
geometry, the exact composition of the gas mixture, and of course the electric 
and the magnetic fields. We used such calculations as a first calibration, but, 
as can be seen with the drift speed, conditions are far from stable in the 
chamber. Moreover, because we use a thin foil as a cathode, the geometry 
knowledge is limited to millimeter precision. These problems, already 
encountered for the BoNuS RTPC calibration (ref), motivated the 
acquisition of specific calibration runs. These are taken with a lower energy 
electron beam (1.20 and 1.27 GeV) to enhance the cross section of the elastic 
process ($e^{4}$He$\rightarrow e^{4}$He). In this process, measurement of only the 
electron kinematics gives the Helium nucleus's kinematic and therefore allows 
to calibrate the drift paths independently of our knowledge of the exact 
conditions in the chamber.

The drift paths' parametrization is possible using identified elastic events 
from our experiment and a simulated copy of them in a GEANT4 
simulation~\cite{GEANT4}. Then by comparing the calculated GEANT4 position of 
the Helium nuclei to the corresponding hits measured in the chamber, we 
can reconstruct the electrons drift path in the RTPC. However, because of the 
magnetic field, the drift paths are not linear. So to perform the extraction, 
we make a first approximation with a linear dependence between the radius of 
emission and the TDC of detection, and then refine our result. As it happens, 
the curvature is minimal and this process converge already on the second 
iteration.

The steps of the extraction procedure are as follows, where we work
in a polar coordinate system ($R$,$\phi$).  The azimuthal difference between
the measured and simulated hit position is $\Delta\phi$.

\begin{figure}[tb]
\centering
%\hspace{-0.95cm}
\includegraphics[scale=0.37]{fig/TdcR_check_p1_10.png}
\caption{$R$ versus TDC distribution for the simulated hits in a single 
$z$-bin. The maximum $R$ is 60 mm (close to the anode) at minimum TDCs 
(the trigger time is 15), and the minimum $R$ is 30 mm (close to the cathode) at 
maximum TDCs ($\sim75$).}
\label{fig:R_correlation}
\end{figure}

\begin{enumerate}

   \item The linear assumption between R and time (TDC) is necessary to link the 
      generated hits from GEANT4 to physical hits. Figure 
      \ref{fig:R_correlation} shows our initial $R$ versus TDC.

   \item For these selected hits, we look at $\Delta\phi$ ($\phi_{sim.}-\phi_{hit\_pad}$)
     versus TDC. We clearly obtain a 
      correlation shown in figure~\ref{fig:1st_pass_delta_phi}. The relationship
      between $\Delta\phi$ and TDC encodes the azimuthal drift of the electrons due to 
      the Lorentz angle.

      \begin{figure}[tb]
         \centering
         \includegraphics[scale=0.37]{fig/TdcPhi_p1_10.png}
         \caption{$\Delta \phi$ versus TDC distribution for the simulated hits 
         with $z$ near 35 mm.}
         \label{fig:1st_pass_delta_phi}
      \end{figure}

   \item Using this first measurement, we can calculate a new correlation 
      between $R$ and TDC. The result, shown in figure~\ref{fig:R_TDC}, is very 
      close to linear and explains the fast convergence of our procedure.

      \begin{figure}[tb]
         \centering
         \includegraphics[scale=0.37]{fig/R_TDC.png}
         \caption{The calculated $R$ versus the time is corrected by the 
         parameters of $\Delta \phi$ from the first pass.}
         \label{fig:R_TDC}
      \end{figure}

   \item In the second pass, we associate simulated helium positions to 
      physical hits according to their calculated correlation and obtain new 
      $\Delta \phi$ distributions.

   \item We finally parametrize the $\Delta\phi$ distributions with a fit as 
      can be seen in figure~\ref{fig:DELTA_PHI_TDC}, which shows the final 
      $\Delta \phi$ versus TDC with a black line representing our drift paths 
      fit.
\end{enumerate}

@@

To verify the stability of the drift paths, this procedure was carried out 
using both the 1.204 GeV data from the beginning of the run period and the 
1.269 GeV data from the end of the run period. We found very similar drift paths
for the two data sets and concluded that any changes in the system only
significantly affected the drift speed.

@@

\end{thebibliography}

>>>>

@@

@@

Measurement of the Deeply Virtual Compton Scattering (DVCS) process
($eH \rightarrow e' H' \gamma$, where $H$ is a hadron) necessitates an upgrade
of this setup.  With a 6 GeV electron beam, the majority of DVCS photons are produced
at very forward angles, where the acceptance of the EC is poor. To extend the detection range,
an inner calorimeter (IC) was built for the E01-113 experiment in 2005.
The IC is constructed from 424 lead-tungstate (PbWO$_{4}$) crystals, covering polar 
angles between 5$^{\circ}$ and 15$^{\circ}$. Each crystal is 16 cm long, corresponding
to 17 radiation lengths. The achieved energy resolution is around 3 to 4$\%$ for photon
energies between 2 GeV and 5 GeV, with an angular resolution between 3 and 5 mrad 
\cite{Hyon-suk}. To protect this detector from the large flux of low energy M{\o}ller 
electrons, a 5~T solenoid was placed around the target to shield the IC. 
In addition, the recoil nuclei of the coherent DVCS on helium was detected 
using a RTPC developed to track low energy nuclear fragments. The solenoid field
serving as the analyzing magnet for the particle tracking in the RTPC. This 
setup was used during a three months experimental run
in 2009 with a longitudinally polarized electron beam of 6.064 GeV incident 
on a gaseous $^{4}$He target.
{\bf \color{red} Add beam current or luminosity.}

@@

- Mention trigger time.

@@

We determine the drift speed using tracks reconstructed in the RTPC. In figure 
\ref{fig:RTPC_signals}, a typical $^{4}$He track is represented (in green). After 
it causes ionization in the drift region, the released electrons (in black) 
drift to the cylindrical detection plane under the effect of the electric field. The 
electrons released close to the cathode take the most time to reach the readout 
pads, but cylindrical symmetry insures they always travel the same 
distance. By identifying the maximum TDC measured, we can infer the drift 
speed of the electrons in the RTPC.\\

@@

To measure the drift speed, we use the time profile of all hits in the 
chamber shown in figure \ref{fig:TDC_profile}. We can clearly observe the 
dropping edge expected from geometrical considerations. We define a value $T_{Max/2}$ at 
which the dropping edge passes half the maximum number of hits in the 
histogram. This value is measure in bins along the 200 mm RTPC's length to take into 
account variations in the electric and magnetic field in the RTPC (see figure 
\ref{fig:RunNumber_61551_TDCmax_Zslice}).

@@

Due to the non perfect experimental conditions, in particular variations in the 
gas mixture\footnote{It was noted during the experiment that leaks appeared
and the gas flow had to be increased in few occasions to compensate for this.}, 
the drift speed changes during the three months long experimental run.  
Figure \ref{fig:Drift_run_number_1} shows the $T_{Max/2}$ values for 
individual runs (approximately 2 hours long). We observe a significant 
variation of the drift speed over time and accounted for it in the drift 
speed used for the track reconstruction.

In summary, we obtain from our calibration a parametrization of the drift speed 
as a function of both position along the beam axis and run number (we did not 
observe any correlation between the effects). These 
functions were extracted for our entire data set and implemented in the 
track reconstructions code.

@@

\subsection{Drift Paths Calibration}

The drift path is the trajectory followed by the electrons released through 
ionization in the gas. Software exists to calculate the drift paths, in particular
MAGBOLTZ~\cite{MAGBOLTZ}, but it requires knowledge of the detector's 
geometry, gas mixture composition, and of course the electric and magnetic 
fields over the whole volume of the detector. We used such calculation as a 
first calibration, but, 
as can be seen with the drift speed, gas composition is far from stable in the 
chamber. Moreover, the 4 $\mu$m foil used as a cathode is easily deformed, such
that we expect the geometry accuracy to be around the millimeter precision. 
These problems, already 
encountered for the BoNuS RTPC calibration~\cite{BONUS-NIM}, motivated the 
acquisition of specific calibration runs. These are taken with a lower energy 
electron beam (1.20 and 1.27 GeV) to enhance the cross section of the elastic 
process ($e^{4}$He$\rightarrow e^{4}$He)
{\bf \color{red} Specify luminosity ?}. In this process, the measurement of
electron kinematics allows to calculate the Helium nucleus kinematic. It is
by comparing this calculated kinematic to the measured one that we fine tune
the drift paths independently of our knowledge of the exact conditions in the 
chamber.

The drift paths are ajusted using a set of identified elastic events 
from our lower beam energy run. Based on the kinematic of the electrons in
these events, we generate the helium nucleus in a GEANT4 
simulation~\cite{GEANT4} of our RTPC. Then we compare the calculated GEANT4 trajectory of 
the Helium nuclei to the hits measured in the chamber.

Because of the 
magnetic field, the drift paths are not linear in the RTPC. So to perform the extraction, 
we make a first approximation with a linear dependence between the radius of 
emission and the time of detection, and then refine our result. As it happens, 
the curvature is minimal and this process converge already on the second 
iteration.

At the end of the extraction procedure, where we work in a polar coordinate 
system ($R$,$\phi$), the azimuthal difference between the detection pad and 
the ionization point ($\Delta\phi$) is extracted as a function of time. 
In figure~\ref{fig:DELTA_PHI_TDC}, we show the resulting data points for one 
bin, where the drift paths is easily identified and eventually fitted for 
implementation in our reconstruction codes.

@@

To verify the stability of the drift paths, this procedure was carried out 
using both the 1.204 GeV data from the beginning of the run period and the 
1.269 GeV data from the end of the run period (shown in blue on 
figure~\ref{fig:Drift_run_number_1}). Strangely, we found very similar drift paths
for the two data sets and concluded that any changes in the system only
significantly affected the drift speed.

@@

\bibitem{BONUS-NIM}
Howard C. Fenker et al., BoNuS: Development and Use of a Radial TPC using Cylindrical GEMs, Nucl. Instrum. Meth. A592, 273-286 (2008)
\end{thebibliography}
